Scalable Clustering of Genotype Information using MapReduce

### 2015-03-09 doc/publications/journal2014/paper.tex
Fixed table

To elaborate, a container is essentially an abstraction of resources, i.e. memory and CPU cores. 
After Hadoop divides a job up into tasks, each task runs in parallel in a container. In addition, each container is independent of other containers.
The resources Hadoop allocates to each container is a predefined value, specified in its configuration. 
% DB reference or version
According to the Cloudera documentation, the general rule of thumb is that one CPU core is allocated to each container, and the available memory per node should be divided equally between each of those containers. 
For example, on a node with 8 CPU cores and 32GB memory, each container should be allocated 1 CPU core and 4GB memory. Effectively allowing a maximum of eight 4GB containers to run in parallel on that node.

@@

To demonstrate the importance of specifying these parameters correctly, we ran the identical jobs, using different Hadoop configurations. 
The \OnePhaseone{} dataset only requires 1.5GB memory per container, as we demonstrated previously.
%Increasing the map container size to 4GB results in the job taking an additional 5 minutes (see Table \ref{Tab:03}).
Increasing the map container size to 4GB results in the job taking an additional 5 minutes (see Table \ref{Tab:03}), because fewer jobs can run in parallel. 
%This larger container size is still too small for the \FullPhasethree{} dataset, which we demonstrated to require 6GB memory per container. 
%Running a job on this dataset with 4GB memory per container results in the job failing.
Attempting to run the larger data set (\FullPhasethree{}) with either 1.5GB or 4GB will result in the job failing, because the feature vector cannot be loaded completely, this requires at least 6GB memory per container and 21h runtime.

One of the main selling points of Hadoop, is that it will run on commodity hardware. 
%For this reason we used a relatively small cluster for our benchmarks. Although our cluster uses VMs on Microsoft Azure, it would be relatively simple for a lab to
%create a similar Hadoop setup in-house. 
To mimic the available resource in a standard resource setting we used only a small VM cluster on Microsoft Azure for the benchmarking in this paper so far.
However, to demonstrate the scaling ability of Hadoop and our MapReduce algorithms, we also ran the clustering jobs on a much larger development Hadoop cluster. 
%As presented in Table \ref{Tab:03}, the \OnePhaseone{} job saw an improvement of approximately 70\%.
%The \FullPhasethree{} example, on the other hand, was 78\% faster.
As presented in Table \ref{Tab:03}, the \OnePhaseone{} job saw an improvement of approximately 70\% when providing 448 cores of 13GB. 
With these resource, the \FullPhasethree{} example, was 78\% faster.

@@

\begin{methods}
\section{Methods}
\subsection*{Datasets}
We used variant call format (VCF) files from phases 1 and 3 of the 1000 Genomes Project.
Each VCF file details the genetic variation (SNPs and indels) for one chromosome, between each individual and a reference genome derived from GRCh37. 
The phase1 and phase3 datasets contain 1,092 and 2,504 individuals, respectively.
Metadata available from the 1000 Genomes Consortium specifies additional attributes for each individual, including population and familial data.\\
We used different subsets of the data for different comparisons. For example, to be able to compare Hadoop with R, we used the VCF file containing chromosome 19 variants (\NinteenPhaseone{}).
To demonstrate scaling with Hadoop, we use subsets that contain more variants, i.e. \SevenPhaseone{} and \FullPhaseone{}, which include chromosome 1-7 and the full genome, respectively.
Table~\ref{datasets} contains an overview of the datasets.

@@

\begin{table*}[!t]
\processtable{Resources allocated and time taken.\label{Tab:02}}
{\begin{tabular}{ll|cc|cc|rr|c}\toprule
Tool & Dataset & \multicolumn{2}{ c |}{Preprocesing (GB)} & \multicolumn{2}{ c |}{Clustering (GB)} & \multicolumn{2}{ c |}{Time} & Accuracy\\
& & map & reduce  & map & reduce  & pre-processing & clustering & \\\midrule
        R &\NinteenPhaseone{} & - & - & - & - & 55m 41s & 7s &0.614\\
        Mahout & \NinteenPhaseone{} & 1536 & 3072 & 1536 & 3072 & 6m 10s & 8m 32s & 0.705\\
        & \OnePhaseone{} & 1536 & 3072 & 1536 & 3072 & 9m 52s & 12m 49s & 0.798\\
        & \SevenPhaseone{} & 1536 & 3072 & 4096 & 4096  & 43m 56s & 3h 28m & 0.813\\
       % & \FullPhaseone{} & 1536 & 3072 & 6144 & 6144  & 85m 29s & 7h 43m & 0.723\\
        & \OnePhasethree{} & 1536 & 3072 & 1536 & 3072 & 23m 27s & 21m 14s & 0.622\\
        & \ThreePhasethree{} & 1536 & 3072 & 4096 & 4096 & 63m 49s & 4h 21m & 0.825\\
        & \FullPhasethree{} & 1536 & 3072 & 6144 & 6144 & 5h 36m & 21h 5m &0.840\\\botrule
\end{tabular}}{The resources allocated to each MapReduce yarn-container and time taken for the job(s) to complete as well as the accuracy measured as \ARI{}.}
\end{table*}

\begin{table*}[!t]
\processtable{Resources allocated and time taken.\label{Tab:03}}
{\begin{tabular}{cl|cc|cc|r}\toprule
Dataset & \multicolumn{2}{ c |}{Available Resources} & \multicolumn{2}{ c |}{Container Size (GB)} & Time\\
& CPU cores & Memory (GB)  & map & reduce & \\\midrule
        \OnePhaseone{} & 24 & 36 & 1536 & 3072 & 12m 49s\\
        \OnePhaseone{} & 24 & 36 & 4096 & 4096  & 17m 47s\\
        \OnePhaseone{} & 448 & 1310 & 1536 & 4096  & 5m 22s\\
        \FullPhasethree{} & 24 & 36 & 6144 & 6144 & 21h 4m\\
        \FullPhasethree{} & 448 & 1310 & 6144 & 6144 & 4h 42m\\\botrule
\end{tabular}}{The resources allocated to each MapReduce yarn-container and time taken for the job(s) to complete.}
\end{table*}

>>>>

To elaborate, a container is essentially an abstraction of resources, i.e. memory and CPU cores. 
After Hadoop divides a job up into tasks, each task runs in parallel in a container. In addition, each container is independent of other containers.
The resources Hadoop allocates to each container is a predefined value, specified in its configuration. 
% DB reference or version
According to the Hortonworks documentation \cite{Horton2013}, the general rule of thumb is that one CPU core is allocated to each container, and the available memory per node should be divided equally between each of those containers. 
For example, on a node with 8 CPU cores and 32GB memory, each container should be allocated 1 CPU core and 4GB memory. Effectively allowing a maximum of eight 4GB containers to run in parallel on that node.

@@

To demonstrate the importance of specifying these parameters correctly, we ran the identical jobs, using different Hadoop configurations. 
The \OnePhaseone{} dataset only requires 1.5GB memory per container, as we demonstrated previously.
%Increasing the map container size to 4GB results in the job taking an additional 5 minutes (see Table \ref{Tab:03}).
Increasing the map container size to 4GB results in the job taking an additional 5 minutes (see Table~\ref{Tab:03}), because fewer jobs can run in parallel. 
%This larger container size is still too small for the \FullPhasethree{} dataset, which we demonstrated to require 6GB memory per container. 
%Running a job on this dataset with 4GB memory per container results in the job failing.
Attempting to run the larger data set (\FullPhasethree{}) with either 1.5GB or 4GB will result in the job failing, because the feature vector cannot be loaded completely, this requires at least 6GB memory per container and 21h runtime.

One of the main selling points of Hadoop, is that it will run on commodity hardware. 
%For this reason we used a relatively small cluster for our benchmarks. Although our cluster uses VMs on Microsoft Azure, it would be relatively simple for a lab to
%create a similar Hadoop setup in-house. 
To mimic the available resource in a standard resource setting we used only a small VM cluster on Microsoft Azure for the benchmarking in this paper so far.
However, to demonstrate the scaling ability of Hadoop and our MapReduce algorithms, we also ran the clustering jobs on a much larger development Hadoop cluster. 
%As presented in Table \ref{Tab:03}, the \OnePhaseone{} job saw an improvement of approximately 70\%.
%The \FullPhasethree{} example, on the other hand, was 78\% faster.
As presented in Table~\ref{Tab:03}, the \OnePhaseone{} job saw an improvement of approximately 70\% when providing 448 cores of 13GB. 
With these resource, the \FullPhasethree{} example, was 78\% faster.

@@

\begin{methods}
\section{Methods}
\subsection*{Datasets}
We used variant call format (VCF) files from phases 1 and 3 of the 1000 Genomes Project.
Each VCF file details the genetic variation (SNPs and indels) for one chromosome, between each individual and a reference genome derived from GRCh37. 
The phase1 and phase3 datasets contain 1,092 and 2,504 individuals, respectively.
Metadata available from the 1000 Genomes Consortium specifies additional attributes for each individual, including population and familial data.\\
We used different subsets of the data for different comparisons. For example, to be able to compare Hadoop with R, we used the VCF file containing chromosome 19 variants (\NinteenPhaseone{}).
To demonstrate scaling with Hadoop, we use subsets that contain more variants, i.e. \SevenPhaseone{} and \FullPhaseone{}, which include chromosome 1-7 and the full genome, respectively.
Table~\ref{Tab:01} contains an overview of the datasets.

@@

\begin{table*}[!t]
\processtable{Resources allocated and time taken.\label{Tab:02}}
{\begin{tabular}{ll|cc|cc|rr|c}\toprule
Tool & Dataset & \multicolumn{2}{ c |}{Preprocesing (MB)} & \multicolumn{2}{ c |}{Clustering (MB)} & \multicolumn{2}{ c |}{Time} & Accuracy\\
& & map & reduce  & map & reduce  & pre-processing & clustering & \\\midrule
        R &\NinteenPhaseone{} & - & - & - & - & 55m 41s & 7s &0.614\\
        Mahout & \NinteenPhaseone{} & 1536 & 3072 & 1536 & 3072 & 6m 10s & 8m 32s & 0.705\\
        & \OnePhaseone{} & 1536 & 3072 & 1536 & 3072 & 9m 52s & 12m 49s & 0.798\\
        & \SevenPhaseone{} & 1536 & 3072 & 4096 & 4096  & 43m 56s & 3h 28m & 0.813\\
       % & \FullPhaseone{} & 1536 & 3072 & 6144 & 6144  & 85m 29s & 7h 43m & 0.723\\
        & \OnePhasethree{} & 1536 & 3072 & 1536 & 3072 & 23m 27s & 21m 14s & 0.622\\
        & \ThreePhasethree{} & 1536 & 3072 & 4096 & 4096 & 63m 49s & 4h 21m & 0.825\\
        & \FullPhasethree{} & 1536 & 3072 & 6144 & 6144 & 5h 36m & 21h 5m &0.840\\\botrule
\end{tabular}}{The resources allocated to each MapReduce yarn-container and time taken for the job(s) to complete as well as the accuracy measured as \ARI{}.}
\end{table*}

\begin{table*}[!t]
\processtable{Resources allocated and time taken.\label{Tab:03}}
{\begin{tabular}{c|cc|cc|c}\toprule
Dataset & \multicolumn{2}{ c |}{Available Resources} & \multicolumn{2}{ c |}{Container Size (MB)} & Time\\
& CPU cores & Memory (GB)  & map & reduce & \\\midrule
        \OnePhaseone{} & 24 & 36 & 1536 & 3072 & 12m 49s\\
        \OnePhaseone{} & 24 & 36 & 4096 & 4096  & 17m 47s\\
        \OnePhaseone{} & 448 & 1310 & 1536 & 4096  & 5m 22s\\
        \FullPhasethree{} & 24 & 36 & 6144 & 6144 & 21h 4m\\
        \FullPhasethree{} & 448 & 1310 & 6144 & 6144 & 4h 42m\\\botrule
\end{tabular}}{The resources allocated to each MapReduce yarn-container and time taken for the job(s) to complete.}
\end{table*}
